---
title: "team.rmd"
author: "Syed Hasan Github @ github.com/shasan2519"
date: "3/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r}
CyDef <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
CyOff <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
bio <- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')
```

# Part 1
```{r}
#1 convert as factor all three sheets
CyOff$Name = as.factor(CyOff$Name)
CyOff$Opponent_Opponent = as.factor(CyOff$Opponent_Opponent)

CyDef$Name = as.factor(CyDef$Name)
CyDef$Opponent_Opponent = as.factor(CyDef$Opponent_Opponent)

bio$Name = as.factor(bio$Name)

#2 convert as numerical/double all sheets
defClean = CyDef %>%
  mutate_if(is.character, as.numeric)

offClean = CyOff %>%
  mutate_if(is.character, as.numeric)

bioClean = bio %>%
  mutate_at("Weight", as.numeric)


#3 convert height to numerical/double only bio
bio <- bio %>%
  separate('Height', c("Feet", "Inches"), 
           sep = "-(?=[^ ]+$)", remove = FALSE) %>%
  mutate_at("Feet", as.numeric) %>%
  mutate_at("Inches", as.numeric)

bio$Height <- bio$Feet * 12 + bio$Inches

bio <- bio %>% select(-Feet, -Inches)

```


```{r}
str(defClean)
str(offClean)
str(bioClean)
```

# Part 2.1
```{r}
#1
defTidy <- defClean %>% 
  pivot_longer(Tackles_Solo:Pass_PB, names_to = 'Statistic',values_to='Value')
head(defTidy)

#2
defTidy %>% 
  ggplot(aes(x = defTidy$Value))+geom_histogram()+facet_wrap(~Statistic)+ggtitle("Defense Stats") +xlab("Occurence")

#3
WestVirginia_KansasState <- defTidy %>% 
  filter(Statistic == 'Tackles_Solo') %>% 
  filter(Opponent_Opponent == 'West Virginia' | Opponent_Opponent == 'Kansas State') %>%
  pivot_wider(names_from='Opponent_Opponent', values_from='Value')

ggplot(data = WestVirginia_KansasState, aes(x = WestVirginia_KansasState$'West Virginia', y = WestVirginia_KansasState$'Kansas State')) +geom_point() +ggtitle("Tackles(individuals)")


```

Rare defensive skills are sacks and turnovers


According to our data Iowa State has a better defense against Kansas State.

# Part 2.2
```{r}
#1
bioTidy <- bioClean %>%
  separate('Hometown', c("Home_City", "Home_State"),sep = ", (?=[^ ]+$)")

bioTidy %>% head(3)

#2
bioTidy %>% group_by(Home_State) %>% summarize(Count = n())

```

Player from each state sorted by state.

# Part 3

Weight and Receivng yards do have a correlation. It is medium and negtive

Brock Purdy is slightly better than others

Overall out of state are equal, but out of state players are slightly better in the defense.
```{r}
#1
part3merge <- right_join(offClean, bioTidy, by = "Name")

#2
part3merge %>% 
  ggplot(aes(x = Weight, y = Receiving_YDS)) +geom_point() +ggtitle("PLayer Stats (weight and receiving)") +xlab("weight") +ylab("yards")

#3
offTidy <- offClean %>%
  pivot_longer(Rushing_ATT:Passing_INT, names_to = 'Off_Statistic',values_to='Off_Value')

team <- full_join(offTidy, defTidy) %>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0))

team$Is_Purdy <- team$Name == 'Purdy, Brock'

team$Is_Purdy <- factor(team$Is_Purdy, labels =  c("Other", "Purdy"))

team <- team %>%
  group_by(Opponent_Opponent, Is_Purdy) %>%
  subset(Off_Statistic == 'Rushing_YDS' | Off_Statistic == 'Receiving_YDS' | Off_Statistic == 'Passing_YDS' |Statistic == 'Tackles_Solo') %>%
  summarize(off_Yards = sum(Off_Value), def_Tackles = sum(Value)) %>%
  pivot_wider(names_from=Is_Purdy, values_from='off_Yards') %>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0)) %>%
  group_by(Opponent_Opponent) %>%
  summarize(Purdy = sum(Purdy), Other = sum(Other), Tackles = sum(def_Tackles))

team %>%
  ggplot(aes(x = Purdy, y = Other)) +geom_point() +ggtitle('Purdy yard vs rest of the teams')

team %>%
  ggplot(aes(x = Purdy, y = Tackles)) +geom_point() +ggtitle('Pudry yards Vs. individual tackles')

#4
bioTidy$Iowan <- bioTidy$Home_State == 'Iowa'

part3merge <- right_join(offClean, bioTidy, by = "Name")

defBioMerge <- right_join(defClean, bioTidy, by = "Name")

part3merge <- part3merge %>% 
  mutate(Rushing_YDS = replace_na(Rushing_YDS, 0)) %>%
  mutate(Passing_YDS = replace_na(Passing_YDS, 0)) %>%
  mutate(Receiving_YDS = replace_na(Receiving_YDS, 0)) %>%
  select(Name, Iowan, Rushing_YDS, Passing_YDS, Receiving_YDS) %>%
  mutate(sum = rowSums(.[3:5]))

part3merge %>%
  group_by(Iowan, Name) %>%
  summarize(Total_YDS = sum(sum)) %>%
  subset(Total_YDS != 0) %>% ggplot(aes(x = Iowan, y = Total_YDS)) +geom_boxplot() +ggtitle("Yard classified by in/out of states")

part3merge %>%
  subset(sum != 0) %>%
  group_by(Iowan) %>%
  summarize(Mean_Player_YDS = sum(sum) / n()) %>%
  print()

part3merge %>%
  subset(sum != 0 & Name != 'Purdy, Brock') %>%
  group_by(Iowan) %>%
  summarize(Mean_Player_YDS = sum(sum) / n(), count = n()) %>%
  print()
```






